Supplement to: Inferring Graphics Programs from Images

### 2017-05-18 supplement.tex
more stuff in the supplement

\newcommand{\Expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\Probability}{\mathds{P}} %{{\rm I\kern-.3em P}}

@@

@@

\section{Neural networks for guiding SMC}

Let $L(\cdot | \cdot):\text{image}^2\to \mathcal{R}$ be our likelihood
function: it takes two images, an observed target image and a
hypothesized program output, and gives the likelihood of the observed
image conditioned on the program output. We want to sample from:
\begin{equation}
\probability [p|x]  \propto L(x | \text{render}(p)) \probability [p]
\end{equation}
where $\probability [p]$ is the prior probability of program $p$, and $x$ is the observed image.

Let $p$ be a program with $L$ lines, which we will write as $p = (p_1,p_2,\cdots,p_L)$. Assume the prior factors into:
\begin{equation}
  \probability [p]\propto \prod_{l\leq L}\probability [p_l]
\end{equation}
Define the distribution $q_L(\cdot)$, which happens to be proportional to the above posterior:
\begin{equation}
  q_L(p_1,p_2,\cdots,p_{L - 1},p_L)\propto q_{L - 1}(p_1,p_2,\cdots,p_{L - 1})\times \frac{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1}))}\times\probability [p_L]
\end{equation}
Now suppose we have some samples from $q_{L - 1}(\cdot)$, and that we
then sample a $p_L$ from a distribution proportional to $\frac{L(x |
  \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x |
  \text{render}(p_1,p_2,\cdots,p_{L - 1}))}\times\probability [p_L]$.
The resulting programs $p$ are distributed according to $q_L$, and so
are also distributed according to $\probability [p|x]$.

How do we sample $p_L$ from a distribution proportional to $\frac{L(x
  | \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x |
  \text{render}(p_1,p_2,\cdots,p_{L - 1}))}\times\probability [p_L]$?
We have a neural network that takes as input the target image $x$ and
the program so far, and produces a distribution over next lines of
code ($p_L$).  We write $\text{NN}(p_L | p_1,\cdots,p_{L - 1};x)$ for
the distribution output by the neural network. So we can sample from NN and then weight the samples by:
\begin{equation}
  w(p_L) = \frac{\probability [p_L]}{\text{NN}(p_L | p_1,\cdots,p_{L - 1};x)}\times \frac{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1}))}
\end{equation}
Then we can resample from these now weighted samples to get a new
population of particles (here programs are particles), where each
program now has $L$ lines instead of $L - 1$.

This procedure can be seen as a particle filter, where each successive
latent variable is another line of code, and the emission
probabilities are successive ratios of likelihoods under $L(\cdot |
\cdot)$.

\textbf{Comments for Dan}. Right now I'm not actually sampling from
the neural network - instead, I enumerate the top few hundred lines of
code suggested by the network, and then weight them by their
likelihoods.
So actually the form of NN is:
\begin{equation}
  \text{NN}(p_L | p_1,\cdots,p_{L - 1};x)\propto \begin{cases}
    1,&\text{ if }p_L \in \text{top hundred neural network proposals}\\
0,&\text{otherwise}.    \end{cases}
\end{equation}
Do you think this is a problem? The neural network puts almost all of its mass on a few guesses.
In order to get the correct line of code I sometimes need to get something like the 50th  top guess,
so I don't want to literally just sample from the distribution suggested by the neural network.

  \begin{algorithm}[tb]
   \caption{Neurally guided SMC}
   \label{guideAlgorithm}
\begin{algorithmic}
  \STATE {\bfseries Input:} Neural network NN, beam size $N$, maximum length $L$, target image $x$
  \STATE {\bfseries Output:} Samples of the program trace
  \STATE Set $B_0 = \{\text{empty program}\}$
  \FOR{$1\leq l\leq L$}
  \FOR{$1\leq n\leq N$}
  \STATE{ $p_n\sim \text{Uniform}(B_{l - 1})$}
  \STATE{ $p'_{n}\sim \text{NN}(\text{render}(p),x)$}
  \STATE{ Define $r_n = p'_n\cdot p_n$}
  \STATE{ Set $\tilde{w}(r_n) = \frac{L(x|r_n)}{L(x|p_n)}\times\frac{\probability [p'_n]}{\probability [p'_n = \text{NN}(\text{render}(p),x)]}$}
  \ENDFOR
  \STATE{ Define $w(p) = \frac{\tilde{w}(p)}{\sum_{p'}\tilde{w}(p')}$}
  \STATE{ Set $B_l$ to be $N$ samples from $r_n$ distributed according to $w(\cdot)$}
  \ENDFOR
  \STATE {\bfseries return} $\{p : p\in B_{l\leq L}, p \text{ is finished}\}$
\end{algorithmic}
  \end{algorithm}

>>>>

\newcommand{\Expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\probability}{\mathds{P}} %{{\rm I\kern-.3em P}}

@@

\section{Neural network architecture}

\subsection{Convolutional network}
The convolutional network takes as input 2 $256\times 256$ images
represented as a $2\times 256\times 256\times$ volume. These are
passed through two layers of convolutions separated by ReLU
nonlinearities and max pooling:
\begin{itemize}
\item Layer 1: 20 $8\times 8$ convolutions, 2 $16\times 4$ convolutions, 2 $4\times 16$ convolutions. Followed by $8\times 8$ pooling with a stride size of 4.
\item Layer 2: 10 $8\times 8$ convolutions. Followed by $4\times 4$ pooling with a stride size of 4.
\end{itemize}
Training takes a little bit less than a day on a Nvidia TitanX GPU.
The network was trained on $10^5$ synthetic examples.

\subsection{Autoregressive decoding of drawing commands}

Given the image features $f$, we predict the first token using logistic regression:
\begin{equation}
  \probability [T_1]\propto W_{T_1}f
\end{equation}
where $W_{T_1}$ is a learned weight matrix.

Subsequent tokens are predicted as:
\begin{equation}
  \probability [T_n|T_{1:(n - 1)}]\propto \text{MLP}_{T_1,n}(I \otimes \bigotimes_{j < n} \text{oneHot}(T_j))
\end{equation}
Thus each token of each drawing primitive has its own learned MLP.
For predicting the coordinates of lines we found that using 32 hidden nodes with sigmoid activations worked well;
for other tokens the MLP's are just logistic regression (no hidden nodes).

\subsection{A learned likelihood surrogate}

Our architecture for
$L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))$ has the
same series of convolutions as the network that predicts the next
drawing command. We train it to predict two scalars: $|T_1 - T_2|$ and
$|T_2 - T_1|$.  These predictions are made using linear regression
from the image features followed by a ReLU nonlinearity; this
nonlinearity makes sense because the predictions can never be negative
but could be arbitrarily large positive numbers.

We train this network by sampling random synthetic scenes for $T_1$,
and then perturbing them in small ways to produce $T_2$.
We minimize the squared loss between the network's prediction and the ground truth symmetric differences.
$T_1$ is rendered in a ``simulated hand drawing'' style which we describe next.

@@

%% \section{Neural networks for guiding SMC}

%% Let $L(\cdot | \cdot):\text{image}^2\to \mathcal{R}$ be our likelihood
%% function: it takes two images, an observed target image and a
%% hypothesized program output, and gives the likelihood of the observed
%% image conditioned on the program output. We want to sample from:
%% \begin{equation}
%% \probability [p|x]  \propto L(x | \text{render}(p)) \probability [p]
%% \end{equation}
%% where $\probability [p]$ is the prior probability of program $p$, and $x$ is the observed image.

%% Let $p$ be a program with $L$ lines, which we will write as $p = (p_1,p_2,\cdots,p_L)$. Assume the prior factors into:
%% \begin{equation}
%%   \probability [p]\propto \prod_{l\leq L}\probability [p_l]
%% \end{equation}
%% Define the distribution $q_L(\cdot)$, which happens to be proportional to the above posterior:
%% \begin{equation}
%%   q_L(p_1,p_2,\cdots,p_{L - 1},p_L)\propto q_{L - 1}(p_1,p_2,\cdots,p_{L - 1})\times \frac{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1}))}\times\probability [p_L]
%% \end{equation}
%% Now suppose we have some samples from $q_{L - 1}(\cdot)$, and that we
%% then sample a $p_L$ from a distribution proportional to $\frac{L(x |
%%   \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x |
%%   \text{render}(p_1,p_2,\cdots,p_{L - 1}))}\times\probability [p_L]$.
%% The resulting programs $p$ are distributed according to $q_L$, and so
%% are also distributed according to $\probability [p|x]$.

%% How do we sample $p_L$ from a distribution proportional to $\frac{L(x
%%   | \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x |
%%   \text{render}(p_1,p_2,\cdots,p_{L - 1}))}\times\probability [p_L]$?
%% We have a neural network that takes as input the target image $x$ and
%% the program so far, and produces a distribution over next lines of
%% code ($p_L$).  We write $\text{NN}(p_L | p_1,\cdots,p_{L - 1};x)$ for
%% the distribution output by the neural network. So we can sample from NN and then weight the samples by:
%% \begin{equation}
%%   w(p_L) = \frac{\probability [p_L]}{\text{NN}(p_L | p_1,\cdots,p_{L - 1};x)}\times \frac{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1}))}
%% \end{equation}
%% Then we can resample from these now weighted samples to get a new
%% population of particles (here programs are particles), where each
%% program now has $L$ lines instead of $L - 1$.

%% This procedure can be seen as a particle filter, where each successive
%% latent variable is another line of code, and the emission
%% probabilities are successive ratios of likelihoods under $L(\cdot |
%% \cdot)$.

%%   \begin{algorithm}[tb]
%%    \caption{Neurally guided SMC}
%%    \label{guideAlgorithm}
%% \begin{algorithmic}
%%   \STATE {\bfseries Input:} Neural network NN, beam size $N$, maximum length $L$, target image $x$
%%   \STATE {\bfseries Output:} Samples of the program trace
%%   \STATE Set $B_0 = \{\text{empty program}\}$
%%   \FOR{$1\leq l\leq L$}
%%   \FOR{$1\leq n\leq N$}
%%   \STATE{ $p_n\sim \text{Uniform}(B_{l - 1})$}
%%   \STATE{ $p'_{n}\sim \text{NN}(\text{render}(p),x)$}
%%   \STATE{ Define $r_n = p'_n\cdot p_n$}
%%   \STATE{ Set $\tilde{w}(r_n) = \frac{L(x|r_n)}{L(x|p_n)}\times\frac{\probability [p'_n]}{\probability [p'_n = \text{NN}(\text{render}(p),x)]}$}
%%   \ENDFOR
%%   \STATE{ Define $w(p) = \frac{\tilde{w}(p)}{\sum_{p'}\tilde{w}(p')}$}
%%   \STATE{ Set $B_l$ to be $N$ samples from $r_n$ distributed according to $w(\cdot)$}
%%   \ENDFOR
%%   \STATE {\bfseries return} $\{p : p\in B_{l\leq L}, p \text{ is finished}\}$
%% \end{algorithmic}
%%   \end{algorithm}
